### 2020-06-16 report/sections/abstract.tex
Report: address some NY comments
Report: new introduction / abstract

We present an end-to-end framework
to write interactive web applications,
communicating over WebSocket transport,
that statically conform to prescribed communication protocols.

We implement a code generation toolchain to
provide developers with TypeScript APIs generated
from a communication protocol specification 
based on multiparty session type theory.

Our toolchain generates APIs that are compatible with
the Node.js runtime and the React.js framework
for server-side and browser-side endpoints respectively.

We evaluate the expressiveness of our API generation solution
using three case studies of protocols found in web-services,
and analyse the overhead of our implementation through 
running micro-benchmarks on a simple binary session protocol.

>>>>

\noindent
Multiparty Session Types (MPST)
is a typing discipline for concurrent
processes that communicate via
message-passing. MPST theory can ensure
communication safety properties and
protocol conformance.
Existing work in session-typed
web development over WebSocket transport
is incompatible with modern web
programming practices and are limited
to supporting communication protocols
that implement the server-centric network
topology.

We address limitations in the current 
state-of-the-art by:
\textbf{(1)} implementing \codegen, a code generation
toolchain for session-typed web development
over WebSocket transport using TypeScript;
and \textbf{(2)} presenting \newtheory, a new
multiparty session type theory that supports
routed communications.

\codegen provides
developers with TypeScript APIs generated
from a communication protocol specification 
based on multiparty session type theory.
Our work is compatible with modern web programming
industrial practices. The generated APIs
build upon TypeScript concurrency primitives,
complement the event-driven style of programming
in full-stack web development,
and are compatible
with the Node.js runtime and the React.js framework
for server-side and browser-side endpoints respectively.
We evaluate the expressiveness of \codegen for modern
web programming
using case studies of protocols found in web-services,
and analyse the overhead of our implementation through 
running micro-benchmarks against a baseline implementation.

\newtheory can express interactions to be routed
via an intermediate participant.
Using \newtheory, 
we propose an approach for supporting 
peer-to-peer communication
between browser-side endpoints through 
routing communication via the server
in a way that avoids excessive serialisation
and preserves communication safety. 
We evaluate the correctness of \newtheory
by proving communication safety properties, such as
deadlock freedom.

### 2020-06-16 report/sections/introduction/contributions.tex
Report: new introduction / abstract

We present an implementation of the MPST framework for developing
interactive web applications over WebSocket transport. 
Related work \citep{PureScript2019, MVU2020, LINKS, Exceptional} requires the 
developer to implement their application using tools that are 
unconventional for web development.
To the best of our knowledge, this is the first work on multiparty session
type-safe web development that provides the same communication safety 
guarantees using familiar industry tools in TypeScript and React.

@@

In \textbf{\cref{chap:codegen}}, 
we introduce the system architecture for our
code generator, \fancyname{SessionTS}. 
The implementation integrates with the Scribble
toolchain to generate handler style APIs that the developer can use to write 
their web applications to guarantee protocol conformance by construction.

We present two code generation strategies for back-end and front-end
TypeScript web development as \fancyname{NodeMPST} and \fancyname{ReactMPST}
in \textbf{\cref{chap:node}} and \textbf{\cref{chap:react}} respectively.

In \textbf{\cref{chap:ext}},
we extend \fancyname{SessionTS} to support advanced
practices commonly found in full-stack web development in a way that
preserves communication safety. We modify the API generation strategies
to support \textit{asynchronous implementations} 
and introduce how to 
gracefully handle errors and \textit{session cancellation} 
in the context of interactive web applications.

@@

\paragraph{\cref{part:evalconcl}}
In \textbf{\cref{chap:eval}}, 
we evaluate our work through 
case studies of communication protocols 
implemented as webservices. Through examples of
multiparty sessions and routed multiparty sessions, 
we demonstrate how the
developer may implement the generated APIs and 
consider the impact of our work
on developer productivity.

@@

\begin{remark}
This project was published as \emph{Generating Interactive WebSocket 
Applications in TypeScript} in Volume 314 of the Electronic Proceedings
in Theoretical Computer Science (EPTCS) \cite{PLACES2020},
in which I am the first and lead author.
The materials in 
\cref{chap:ext,chap:p2p,chap:theory,chap:impl,chap:eval}
are developed exclusively in this project by myself.
\end{remark}

>>>>

We present an implementation of the MPST framework for developing
interactive web applications over WebSocket transport.
We also present a theory of routed multiparty session types,
which forms the basis for supporting arbitrary communication structures
over a server-centric WebSocket-based network topology.
Related work \citep{PureScript2019, MVU2020, LINKS, Exceptional} 
requires the 
developer to implement their application using tools that are 
unconventional for web development, 
and only support communication protocols that adhere to the
WebSocket network structure.
To the best of our knowledge, this is the first work on multiparty session
type-safe web development that provides the same communication safety 
guarantees using familiar industry tools in TypeScript and React,
and supports peer-to-peer browser interactions over WebSocket transport.

@@

In \textbf{\cref{chap:codegen}}, 
we introduce the system architecture for
\codegen. 
The implementation integrates with the Scribble
toolchain to generate handler style APIs that the developer can use to write 
their web applications to guarantee protocol conformance by construction.

We present two code generation strategies for back-end and front-end
TypeScript web development as \nodecodegen and \reactcodegen
in \textbf{\cref{chap:node}} and \textbf{\cref{chap:react}} respectively.

In \textbf{\cref{chap:ext}},
we extend \codegen to support advanced
practices commonly found in full-stack web development in a way that
preserves communication safety. 
We modify the API generation strategies
to support \textit{asynchronous implementations} 
and introduce how to 
gracefully handle errors and \textit{session cancellation} 
in the context of interactive web applications.

@@

\paragraph{\cref{part:evalconcl}}
In \textbf{\cref{chap:eval}}, 
we evaluate our work through 
case studies of communication protocols 
implemented as webservices. Through examples of
multiparty sessions and routed multiparty sessions, 
we demonstrate how the
developer may implement the generated APIs and 
consider the compatibility of our work with respect to
modern web programming practices.
We perform benchmarks on a web application built using our
generated APIs against a baseline implementation to evaluate
the overhead of our work.

@@

\begin{remark}
This project was published as \emph{Generating Interactive WebSocket 
Applications in TypeScript} in Volume 314 of the Electronic Proceedings
in Theoretical Computer Science (EPTCS) \cite{PLACES2020},
in which I am the first and lead author.

The materials in \cref{chap:codegen,chap:node,chap:react}
are revised from \cite{PLACES2020}.
The materials in 
\cref{chap:ext,chap:p2p,chap:theory,chap:impl,chap:eval}
are developed exclusively in this project by myself.
\end{remark}

### 2020-06-16 report/sections/introduction/objectives.tex
Report: new introduction / abstract

\subparagraph{S1}
We implement \codegen, an end-to-end framework
for writing full-stack \textit{TypeScript} applications that
\textbf{statically} conform to a communication protocol.

@@

\subparagraph{S2}
We present \newtheory, an extension of the canonical MPST
theory that supports \textit{routed communication}.

>>>>

\subparagraph{S1: \codegen}
We implement an end-to-end framework
for writing full-stack \textit{TypeScript} applications that
\textbf{statically} conform to a communication protocol.

@@

\subparagraph{S2: \newtheory}
We present an extension of the canonical MPST
theory that supports \textit{routed communication}.
